---
title: "Statistical report: Outcomes of stable angina patients with non-obstructive coronary artery disease"
subtitle: DRAFT
date: "`r Sys.Date()`"
cover: static/ki_logo_vert_rgb.png
editor: source
format:
  pdf:
    documentclass: scrartcl
    template-partials:
      - "static/before-body.tex"
      - "static/_titlepage.tex"
    include-in-header: 
      - "static/in-header.tex"
    toc: true
    toc-depth: 3
    toc-title: Contents
    lof: true
    lot: true
    cap-location: top
    number-sections: true
    colorlinks: false
    keep-tex: false
bibliography: static/references.bib
csl: static/bmj.csl
link-citations: true
link-bibliography: true
nocite: '@*'
knitr:
  opts_chunk: 
    dev: cairo_pdf
    collapse: true
    comment: "" 
    echo: FALSE
    include: TRUE
    warning: FALSE
    message: FALSE
    fig.pos: "H"
    fig.path: "../output/figs/"
    fig.height: 6
    fig.width: 10
    R.options:
      knitr.graphics.auto_pdf: true
---

```{r}
#| label: set-up-load-data
options(knitr.kable.NA = "")

# load packages, globals and project specific functions
source(here::here("setup/setup.R"))

# load data
load(here("data/clean-data/sdata.RData"))

# load workbook to write tables to Excel
wb <- loadWorkbook(here("output/tabs/tables.xlsx"))
sheets <- names(wb)

# load pptx file with figs
figs <- officer::read_pptx(path = here::here("output/figs/figs.pptx"))
pptx <- TRUE
```           

\newpage

# Data

## Data sources

SWEDEHEART - SCAAR is an ongoing nationwide quality register. 
SCAAR was linked with national Swedish registries through the Swedish personal identity number (PIN)[@pin]: 

- The National Prescribed Drug Register (The National Board of Health and Welfare) for information on medications
- The National Patient Register (The National Board of Health and Welfare) for additional comorbidities and outcomes
- The Cause of Death Register[@dors] (The National Board of Health and Welfare) for outcomes death
- Longitudinal integrated database for health insurance and labour market studies (LISA)[@lisa] (Statistics Sweden) for socio-economic factors and selection of controls

## Inclusion/exclusion criteria

```{r}
#| label: tbl-flow
#| tbl-cap: Information for flowchart
#| tbl-pos: "H"

default_kable(flow)
```

First patient in: `r min(sdata$INTERDAT, na.rm = T)` and last patient in: `r max(sdata$INTERDAT, na.rm = T)` 
from approximately `r sdata %>% distinct(CENTREID) %>% count() %>% pull(n)` enrolling centers. 

Baseline is set to the intervention date + `r global_indexplus` days. The `r global_indexplus` days is in order to take into account that medications can be collected from the pharmacy up to `r global_indexplus` days after the intervention and to avoid that the outcomes are baseline events. 

```{r}
#| label: tbl-ncontrols
#| tbl-cap: Number of controls
#| tbl-pos: "H"

ncontrols <- sdata %>%
  filter(case == "ANOCA") %>%
  count(ncontrols) %>%
  select(n, ncontrols)
names(ncontrols) <- c("N cases with", "N controls")

default_kable(ncontrols, scale_down = F)
```

## Definitions

### Controls

The original control population was matched to another population by year of birth, sex and county of residence assuming no previous indication (ANGE VILKEN INDIKATION). The re-matching could therefore cause a bias since the risk of INDIKATION will be smaller in this pool of persons compared to the general population. 

For this study the 2 controls / case were re-matched by year of birth, sex and county of residence and without prior registration in SCAAR or indication according to the National Patient Register (Procedure codes FNA-H or diagnosis codes I01.2, I09.I21–22 (MI), I20.0 (IAP), I01.2, I09.0, I40, I41, I42.3, I51.4 (Myocarditis). The control can however experience either of these conditions later during follow-up and thereby theoretically be included twice in the analyses, partly as a control for another case and partly, later on, as a case. 

### Variables

{{< include src/vars.qmd >}}

### Time-dependent covariates (medications)

A number of medications were included as time-dependent covariates in the cox models. The medications were derived from The National Prescribed Drug Register. If a person not on the medication at baseline purchased any of the respective medications during follow-up the person is from that time considered to be on the respective medication. A prescription in Sweden usually lasts 3 months and to make allowance for non-compliance, pill-stocking etc. a time-limit of 5 months was set were if a new prescription had not been purchased from the last dispensation within this time-frame the person was defined as not on the medication. If a new prescription is filled at a later date the person is yet again to be considered as on the respective medication. 

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with care.

## Missing data

```{r}
missedu <- sdata %>%
  mutate(isna = ynfac(if_else(is.na(scb_education), 1, 0))) %>%
  count(isna, .drop = F) %>%
  mutate(np = paste0(n, " (", fn(n / sum(n) * 100, 1), "%)")) %>%
  filter(isna == "Yes") %>%
  pull(np)
missmar <- sdata %>%
  mutate(isna = ynfac(if_else(is.na(scb_maritalstatus), 1, 0))) %>%
  count(isna, .drop = F) %>%
  mutate(np = paste0(n, " (", fn(n / sum(n) * 100, 1), "%)")) %>%
  filter(isna == "Yes") %>%
  pull(np)
missinc <- sdata %>%
  mutate(isna = ynfac(if_else(is.na(scb_dispincome), 1, 0))) %>%
  count(isna, .drop = F) %>%
  mutate(np = paste0(n, " (", fn(n / sum(n) * 100, 1), "%)")) %>%
  filter(isna == "Yes") %>%
  pull(np)
```

In the analyses where cases were compared to their matched controls absence of a prescription/ICD/procedure code was considered as No, the proportion of 
missing data was low and only occurred for education in `r missedu` persons, for marital status in `r missmar` persons and for income in `r missinc` persons. 
Due to the low % of missing data no multiple imputation was performed but type (for education and marital status) and median within year 
(disposable income) were imputed when used in the multivariable models. NOTE TO SELF: Check if any missing marital and income!!!!!

For the analyses where only cases were investigated (where the variables of interest also where taken from SCAAR) 
missing data was imputed with multiple imputation (n = 10) using mice [@mice] 
and Rubin’s rules was used for combining estimates and standard errors across imputed data sets. 
Variables included in the model are indicated in @tbl-base. The primary outcome, `r outvars %>% filter(primary) %>% pull(name)`, was included as the Nelson-Aalen estimator 
in the model but itself not imputed since it contained no missing values.

## Subgroups

The case/control analyses were also performed according to sex, age-groups and year of inclusion. 

## Baseline characteristics

Baseline characteristics are presented by cases/controls. 
Categorical variables are presented with n and percentage and tested for 
differences between groups with the chi-square test. 
Continuous variables are presented with median [first quartile-third quartile] and tested for 
differences between groups with the Kruskal-Wallis test. 

## Outcomes

The following outcomes are considered: 

- `r paste0(outvars$name[1], " (", outvars$shortname[1], ") (primary outcome)")`
- `r paste0(outvars$name[2], " (", outvars$shortname[2], ")")`
- `r paste0(outvars$name[3], " (", outvars$shortname[3], ")")`
- `r paste0(outvars$name[4], " (", outvars$shortname[4], ")")`
- `r paste0(outvars$name[5], " (", outvars$shortname[5], ")")`
- `r paste0(outvars$name[6], " (", outvars$shortname[6], ")")`
- `r paste0(outvars$name[7])`
- `r paste0(outvars$name[8], " (", outvars$shortname[8], ")")`

Data were censored at `r global_endfollowup`, emigration from Sweden or at death if itself not an event. 

### Association between outcomes and cases/controls

Time to first events were presented with cumulative incidence curves by case/control. 

Cox proportional hazards regressions were used to model the time to first event including the matched pairs as a frailty term in the model.
Crude and adjusted analyses were performed where adjustment was performed for 
variables indicated in @tbl-base. And, by design due to matching, also for age, sex and county of residence. 

Further, as a consistency analysis, Lipid lowering, glucose lowering, anticoagulation, Antiplatelet therapy, bbl, rasi/arni, ccb were modelled as time-dependepent covariates in the model. 

As a consistency analysis for the primary outcome, a sub-distributional hazards model[@fg] was used, 
where deaths not defined as an event were treated as competing events (not performed for the subgroups).

### Association between primary outcome and basline characteristics for cases

For the ANOCA cases only cox proportional hazards regressions were used to model 
the association between partly the primary outcome and partly revascularization and the variables indicated in @tbl-base. 
Note that the variables collected in SCAAR are at the date of intervention whereas the other variables are from the date of intervention + `r global_indexplus` days. 

### Assumptions

Note to self: TO BE CHECKED!!!! The proportional hazards assumption was investigated using the scaled Schoenfeld 
residuals leading to that `r paste0(stratavars, collapse =", ")` were stratified for in the models.  
Possible outliers were visually inspected by plotting the dfbetas. 

\clearpage

# Results

```{r}
med <- sdata %>%
  summarise(
    med = fn(median(scb_age), dig = 0),
    q1 = fn(quantile(scb_age, probs = 0.25), dig = 0),
    q3 = fn(quantile(scb_age, probs = 0.75), dig = 0)
  ) %>%
  mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
  pull(out)

percfem <- sdata %>%
  count(scb_sex) %>%
  mutate(perc = fn(n / sum(n) * 100, 0)) %>%
  filter(scb_sex == "Female") %>%
  pull(perc)
```

The median (q1-q3) age is `r med` and `r percfem`% females. 

## Baseline characteristics

{{< include src/basetab.qmd >}}

\clearpage

## Association between outcomes and cases/controls

```{r}
medsurv <- sdata %>%
  summarise(
    med = fn(median(sos_outtime_death / 365.25), dig = 0),
    min = fn(min(sos_outtime_death / 365.25), dig = 0),
    max = fn(max(sos_outtime_death / 365.25), dig = 0)
  ) %>%
  mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
  pull(out)

sumsurv <- sdata %>%
  summarise(sumpy = fn(sum(sos_outtime_death) / 365.25, dig = 0)) %>%
  pull(sumpy)

medsurv_case <- sdata %>%
  filter(case == "ANOCA") %>%
  summarise(
    med = fn(median(sos_outtime_death / 365.25), dig = 0),
    min = fn(min(sos_outtime_death / 365.25), dig = 0),
    max = fn(max(sos_outtime_death / 365.25), dig = 0)
  ) %>%
  mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
  pull(out)

sumsurv_case <- sdata %>%
  filter(case == "Case") %>%
  summarise(sumpy = fn(sum(sos_outtime_death) / 365.25, dig = 0)) %>%
  pull(sumpy)
```

The median (min-max) follow-up is 
`r medsurv` years for a total of `r sumsurv` patient-years of follow-up for the cases and controls and 
`r medsurv_case` years for a total of `r sumsurv_case` patient-years of follow-up for the cases only.

{{< include src/outcomes.qmd >}}

\clearpage

{{< include src/kmfigs.qmd >}}

\clearpage

## Association between primary outcome and baseline characteristics for ANOCA only

{{< include src/predictors.qmd >}}

\clearpage

# Reproducibility

## R code

The R code for all data handling and statistical analyses are found: 

https://github.com/KIHeartFailure/scaar-stableangina. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

# References

::: {#refs}
:::
